Path Model Temperature Comparisons

### 2016-07-15 path-temp-compare.Rmd
Revise model formulae and add regression residual plots.

```{r load data, echo=F}
load("climate-path-model-data.RData")
library(piecewiseSEM); library(dplyr); library(lme4); library(lmerTest)
oakplots.sub$umca_density_m2 <- oakplots.sub$tot_bay / (15*15)
oakplots.sub$umca_density_m2.sqrt <- sqrt(oakplots.sub$umca_density_m2)
```

@@

```{r rescale variables}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- scale(oakplots.sub$tot_bay)
oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)
oakplots.sub$umca_ba.log.scl <- scale(oakplots.sub$umca_ba.log)
oakplots.sub$tot_lfct.scl <- scale(oakplots.sub$tot_lfct)
oakplots.sub$tot_lfct.log.scl <- scale(oakplots.sub$tot_lfct.log)
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- scale(oakplots.sub$hrs_abv25ds)
oakplots.sub$avg_tmax_ds.scl <- scale(oakplots.sub$avg_tmax_ds)
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- scale(oakplots.sub$avg_hrsblw14_wet_tminus1)
oakplots.sub$hrs1422_wet_tminus1.log.scl <- scale(oakplots.sub$hrs1422_wet_tminus1.log)
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1.log)

@@

Using data on oak infection at the plot-level:
```{r model pieces}
names(oakplots.sub)
oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)

model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),

      umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),

@@

      avgtmax_ds <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

      umca_density <- lm(tot_bay.log.scl ~ twi15m.log.scl, data = oakplots.sub),

@@

      avgtmax_ds <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)

# Model set reduced to plots that have oaks & at least 1 bay laurel
oak_and_bay_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oak_and_bay_plots),
      #umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),
      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      hrs_abv25 <- lmer(hrs_abv25ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
)

# Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
                        "tot_bay.log ~~ H.2005"))

sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
                          "tot_bay.log ~~ H.2005"))

# Model set using average daily maximum temperature during the dry season.
model_avgtmax_ds <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),

>>>>

```{r load data, echo=F}
load("climate-path-model-data.RData")
library(piecewiseSEM); library(dplyr); library(lme4); library(lmerTest)

# Calculate oak density variable as stems per hectare 
oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
```

@@

```{r rescale variables, echo=F}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- scale(oakplots.sub$tot_bay)
oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
oakplots.sub$umca_density_ha.scl <- scale(oakplots.sub$umca_density_ha)
oakplots.sub$umca_density_ha.log.scl <- scale(oakplots.sub$umca_density_ha.log)
oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)
oakplots.sub$umca_ba.log.scl <- scale(oakplots.sub$umca_ba.log)
oakplots.sub$tot_lfct.scl <- scale(oakplots.sub$tot_lfct)
oakplots.sub$tot_lfct.log.scl <- scale(oakplots.sub$tot_lfct.log)
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- scale(oakplots.sub$hrs_abv25ds)
oakplots.sub$dys_abv25ds.scl <- scale(oakplots.sub$dys_abv25ds)
oakplots.sub$avg_tmax_ds.scl <- scale(oakplots.sub$avg_tmax_ds)
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- scale(oakplots.sub$avg_hrsblw14_wet_tminus1)
oakplots.sub$hrs1422_wet_tminus1.log.scl <- scale(oakplots.sub$hrs1422_wet_tminus1.log)
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1.log)

@@

Using data on oak infection at the plot-level:
```{r hours >25 path model all oak plots unscaled, echo=F}
model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.sub, sample_year == 2005)),

      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

@@

      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one. Each variable is first  centered by subtracting its mean and then scaled by dividing by its standard deviation.
```{r hours >25 path model all oak plots scaled}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

@@

      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

Examining the residual plots of each piece of the path model
```{r shannons residuals plot}
par(mfrow=c(2,2))
plot(shannons, main = "OLS regression of diversity ~ wetness index")
```

```{r bay laurel density residuals plot}
plot(umca_density, main = "LMM of bay laurel density")
```

```{r symptomatic leaf count residuals plot}
plot(slc, main = "LMM of symptomatic leaf count")
```

```{r wet-hours residuals plot}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year")
```

```{r dry-season hours residuals plot}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season")
```

```{r oak infection residuals plot}
plot(oak_infection, main = "Binomial GLMM of oak infection")
```

Model set reduced to plots that have oaks & at least 1 bay laurel
```{r path model with oak plots having >0 bay laurel}
oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)
oak_and_bay_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oak_and_bay_plots),
      #umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),
      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      hrs_abv25 <- lmer(hrs_abv25ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
)
```

```{r fit and compare all oak plots vs >0 bay laurel oak plots}
# Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
                        "tot_bay.log ~~ H.2005"))
```

Get the path coeffcients for the fitted path models
```{r get path coefficients}
sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))

sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
                          "tot_bay.log ~~ H.2005"))
```

```{r path models using avg dry season tmax}
# Model set using average daily maximum temperature during the dry season.
model_avgtmax_ds <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),
